Composable Generalized Consensus

### 2016-01-10 paper/intro.tex
Working on presenting CGC.

An SMR module is an SMR algorithm augmented with \emph{composition interface}
and satisfying our correctness condition, called Composable Generalized
Consensus (CGC). The composition interface consists of an \emph{abort interface}
for aborting an SMR module's execution and passing the baton to another
SMR module and with a matching \emph{init interface} for taking over the
execution of an aborting SMR module. Two SMR modules are composed by passing
the outputs produced by the abort interface of the first SMR module, called
abort values, as inputs to the init interface of the second SMR module,
called init values in the context of the second SMR module. A SMR module
must satisfy a correctness condition that we call Composable Generalized
Consensus (abbreviated CGC). Any correct SMR module is a safe Generalized
Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR
algorithm), but the progress guarantees may vary between SMR modules: CGC only
requires a SMR module to make progress executing commands or to abort its
execution. CGC also relates the init values, learned values (learned values
represent the local state-machine state of individual replicas), and abort
values in order for our composition theorem to hold.
%Note that Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can offer important performance advantages in practice.

>>>>

An SMR module is an SMR algorithm augmented with a \emph{composition interface}
and satisfying our correctness condition, called Composable Generalized
Consensus (CGC). The composition interface consists of an \emph{abort interface}
for aborting an SMR module's execution and passing the baton to another
SMR module and with a matching \emph{init interface} for taking over the
execution of an aborting SMR module. Two SMR modules are composed by passing
the outputs produced by the abort interface of the first SMR module, called
abort values, as inputs to the init interface of the second SMR module,
called init values in the context of the second SMR module. A SMR module
must satisfy a correctness condition that we call Composable Generalized
Consensus (abbreviated CGC). Any correct SMR module is a safe Generalized
Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR
algorithm), but the progress guarantees may vary between SMR modules: CGC only
requires a SMR module to make progress executing commands or to abort its
execution. CGC also relates the init values, learned values (learned values
represent the local state-machine state of individual replicas), and abort
values in order for our composition theorem to hold.
%Note that Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can offer important performance advantages in practice.

### 2016-01-10 paper/related.tex
Work on the intro, mainly.
Working on presenting CGC.

Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and allows its implementation to optimize the execution of commuting commands. In cases in which many commands commute, which may be frequent in practice, Generalized Consensus algorithms can have a significant performance advantage over traditional consensus algorithms.

Making SMR algorithms efficient has been a long-standing
challenge and a multitude of algorithms have been proposed, each best suited to a particular use case.
Examples include Paxos~\cite{lamport2001paxos}, Fast
Paxos~\cite{Lamport06FastPaxos}, Disk Paxos~\cite{GafniLamport03DiskPaxos},
Generalized Paxos~\cite{Lamport05GeneralizeConsensus},
FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus}, Chain
Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, Ring
Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
Egalitarian Paxos
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
Multi-Coordinated Paxos~\cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
Cheap Paxos~\cite{LamportMassa04CheapPaxos}, Vertical Paxos
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting}, Mencius
\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
Fast Mencius~\cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
the Chandra-Toueg algorithm
\cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
Ben-Or algorithm~\cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
and Raft~\cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

Our work is closely related to the Abstract
framework~\cite{GuerraouiETAL10Next700BftProtocols}
and the Speculative Linearizability
framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}. The
Abstract framework pioneered the idea of decomposing SMR algorithms in
independent modules and experimentally showed that the resulting composite
algorithms outperformed state-of-the-art monolithic algorithms. Compared
to our work, the Abstract framework is not precisely formalized and lacks
tools supporting the development of correct algorithms in the framework,
which are direly needed given the notorious complexity of SMR algorithms.
Moreover, the Abstract framework does not allow optimizing for the execution
of commuting commands. However, the Abstract framework supports byzantine
fault-tolerant~\cite{LamportShostakPease82ByzantineGeneralsProblem} SMR
algorithms, which we do not.

The Speculative Linearizability framework was a
first attempts by the authors to formalize the Abstract framework, provide
formal proof support in Isabelle/HOL~\cite{NipkowPaulsonWenzel02IsabelleHOL},
and extend the Abstract framework to support algorithms that optimize
the execution of commuting commands by requiring Speculatively
Linearizable modules to refine Abstract modules (under a forward/backward
simulation~\cite{LynchVaandrager95ForwardBackwardSimulationsIUntimedSystems})
modulo an equivalence relation on histories of commands that reflects the
commutativity relation among commands. However, the simulation technique led to
complex correctness statements that were hard to scale beyond the examples of
the paper, and no realistic protocol was built in the framework. The present
work addresses the shortcoming of the Speculatively Linearizable framework by
using the notion of command-structure set introduced for the definition of
Generalized Consensus~\cite{Lamport05GeneralizeConsensus}. Moreover, we show how
several well known and practical algorithms can be cast and optimized in our
framework, providing a set of TLA+ specifications that ease the design and model
checking of algorithms based on quorum systems.

TODO\@: expand the following.

Generalized Consensus is an agreement problem relaxing the ordering constraints
imposed on commuting commands and Generalized Paxos~\cite{Lamport05GeneralizeConsensus} is an algorithm, similar
in structure to Fast Paxos, which takes advantage of the relaxed ordering
constraint to process commuting commands in two message delays, when three are
needed in the Classic Paxos algorithm.

@@

Vertical Paxos algorithms are SMR algorithms whose set of replica servers can
be reconfigured without stopping to process new commands. Crashed servers can
therefore be replaced by new servers, ensuring the long-term availability of a
replicated service. Vertical Paxos algorithms use an external reconfiguration
master, itself implemented as a replicated state machine, which determines
the servers that should be removed from or added to the configuration. The
role of the reconfiguration master is similar to the role of the scheduling
policy in our framework. In contrast to our work, Vertical Paxos does not define
an interface and correctness condition for encapsulating the reconfigurable
instances of SMR algorithms and therefore does not allow the reuse of existing
algorithms and the incremental development of SMR optimizations.

>>>>

% Abstract framework
Our work is closely related to the Abstract
framework~\cite{GuerraouiETAL10Next700BftProtocols}
and the Speculative Linearizability
framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}. The
Abstract framework pioneered the idea of decomposing SMR algorithms in
independent modules and experimentally showed that the resulting composite
algorithms outperformed state-of-the-art monolithic algorithms. Compared
to our work, the Abstract framework is not precisely formalized and lacks
tools supporting the development of correct algorithms in the framework,
which are direly needed given the notorious complexity of SMR algorithms.
Moreover, the Abstract framework does not allow optimizing for the execution
of commuting commands. However, the Abstract framework supports byzantine
fault-tolerant~\cite{LamportShostakPease82ByzantineGeneralsProblem} SMR
algorithms, which we do not.

% Speclin
The Speculative Linearizability framework was a
first attempts by the authors to formalize the Abstract framework, provide
formal proof support in Isabelle/HOL~\cite{NipkowPaulsonWenzel02IsabelleHOL},
and extend the Abstract framework to support algorithms that optimize
the execution of commuting commands by requiring Speculatively
Linearizable modules to refine Abstract modules (under a forward/backward
simulation~\cite{LynchVaandrager95ForwardBackwardSimulationsIUntimedSystems})
modulo an equivalence relation on histories of commands that reflects the
commutativity relation among commands. However, the simulation technique led to
complex correctness statements that were hard to scale beyond the examples of
the paper, and no realistic protocol was built in the framework. The present
work addresses the shortcoming of the Speculatively Linearizable framework by
using the notion of command-structure set introduced for the definition of
Generalized Consensus~\cite{Lamport05GeneralizeConsensus}. Moreover, we show how
several well known and practical algorithms can be cast and optimized in our
framework, providing a set of TLA+ specifications that ease the design and model
checking of algorithms based on quorum systems.

% Generalized Paxos
Generalized Consensus is an agreement problem relaxing the ordering constraints
imposed on commuting commands and Generalized Paxos~\cite{Lamport05GeneralizeConsensus} is an algorithm, similar
in structure to Fast Paxos, which takes advantage of the relaxed ordering
constraint to process commuting commands in two message delays, when three are
needed in the Classic Paxos algorithm.

@@

% Vertical Paxos
Vertical Paxos algorithms are SMR algorithms whose set of replica servers can
be reconfigured without stopping to process new commands. Crashed servers can
therefore be replaced by new servers, ensuring the long-term availability of a
replicated service. Vertical Paxos algorithms use an external reconfiguration
master, itself implemented as a replicated state machine, which determines
the servers that should be removed from or added to the configuration. The
role of the reconfiguration master is similar to the role of the scheduling
policy in our framework. In contrast to our work, Vertical Paxos does not define
an interface and correctness condition for encapsulating the reconfigurable
instances of SMR algorithms and therefore does not allow the reuse of existing
algorithms and the incremental development of SMR optimizations.
